---
title: "Greta and Causact: Some Examples"
author: "Peter Rabinovitch"
date: "`r Sys.time()`"
output: github_document
always_allow_html: true
---


# Introduction

About a year ago I wrote a [post](https://prabinov42.github.io/posts/greta_and_causact/) abut using the *causact* R library, which uses the *greta* library, which uses *TensorFlow* to estimate Bayesian models.  

As I have had spare time I have converted several of the [greta examples](https://greta-stats.org/articles/example_models.html) into causact code. I plan to do more as time allows....including cleaning it up a bit with more explanations.

Each example follows the same pattern: duplicate the greta model from the examples page, run it and pull out some results. Then repeat the model in causact, display the graph (love these), and pull out some results.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(bayesplot)
library(greta)
library(causact)
```

# Example 1: Linear regression

```{r}
# A simple, one-variable Bayesian linear regression model using the attitude data

# greta
# variables & priors
data("attitude")
int <- normal(0, 10)
coef <- normal(0, 10)
sd <- cauchy(0, 3, truncation = c(0, Inf))
# linear predictor
mu <- int + coef * attitude$complaints
# observation model
distribution(attitude$rating) <- normal(mu, sd)
gretaModel <- model(int, coef, sd, mu) # MODEL
draws <- mcmc(gretaModel)
tidyDrawsDF <- bind_rows(
  draws$`11` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 1) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value"),
  draws$`12` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 2) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value"),
  draws$`13` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 3) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value"),
  draws$`14` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 4) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value")
)

tidyDrawsDF %>%
  pull(param) %>%
  unique()
tidyDrawsDF %>%
  filter(param == "coef") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>%
  filter(param == "int") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>%
  filter(param == "sd") %>%
  pull(value) %>%
  mean()

# causact
df <- attitude %>% select(rating, complaints)
graph <- dag_create() %>%
  dag_node("rating", "rating", rhs = normal(mu, sd), data = df$rating) %>%
  dag_node("mu", "mu", rhs = int + coef * complaints, child = "rating") %>%
  dag_node("sd", "sd", rhs = cauchy(0, 3, truncation = c(0, Inf)), child = "rating") %>%
  dag_node("int", "int", rhs = normal(0, 10), child = "mu") %>%
  dag_node("coef", "coef", rhs = normal(0, 10), child = "mu") %>%
  dag_node("complaints", "complaints", data = df$complaints, child = "mu") %>%
  dag_plate("Observations", "i", nodeLabels = c("rating", "complaints", "mu"))
graph %>% dag_render()
set.seed(2022)
drawsDF <- graph %>% dag_greta(mcmc = TRUE)
drawsDF %>% dagp_plot()
```

# Example 2: Multiple linear regression

```{r}
# A multi-variable Bayesian linear regression model using the attitude data
# greta
data(attitude)
design <- as.matrix(attitude[, 2:7])
int <- normal(0, 10)
coefs <- normal(0, 10, dim = ncol(design))
sd <- cauchy(0, 3, truncation = c(0, Inf))
# matrix multiplication is more efficient than multiplying the coefficients
# separately
mu <- int + design %*% coefs
distribution(attitude$rating) <- normal(mu, sd)
gretaModel <- model(int, coefs, sd) # MODEL
draws <- mcmc(gretaModel)
tidyDrawsDF <- replaceLabels(draws) %>%
  as.matrix() %>%
  as_tibble() %>%
  addPriorGroups()
tidyDrawsDF %>% glimpse()
tidyDrawsDF %>%
  pull(param) %>%
  unique()
tidyDrawsDF %>%
  filter(param == "coefs[1,1]") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>%
  filter(param == "coefs[6,1]") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>%
  filter(param == "int") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>%
  filter(param == "sd") %>%
  pull(value) %>%
  mean()

tidyDrawsDF <- bind_rows(
  draws$`11` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 1) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value"),
  draws$`12` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 2) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value"),
  draws$`13` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 3) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value"),
  draws$`14` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 4) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value")
)

# causact
design <- as.matrix(attitude[, 2:7])
graph <- dag_create() %>%
  dag_node("rating", "rating", rhs = normal(mu, sd), data = df$rating) %>%
  dag_node("mu", "mu", rhs = int + design %*% coefs, child = "rating") %>%
  dag_node("sd", "sd", rhs = cauchy(0, 3, truncation = c(0, Inf)), child = "rating") %>%
  dag_node("int", "int", rhs = normal(0, 10), child = "mu") %>%
  dag_node("coefs", "coefs", rhs = normal(0, 10, dim = ncol(design)), child = "mu") %>%
  dag_node("design", "design", data = design, child = "mu", keepAsDF = TRUE) %>%
  dag_plate("Observations", "i", nodeLabels = c("rating", "design", "mu"))
graph %>% dag_render()
set.seed(2022)
drawsDF <- graph %>% dag_greta(mcmc = TRUE)
drawsDF %>% dagp_plot()
#
# #or
# graph <- dag_create() %>%
#   dag_node("rating", "rating", rhs = normal(mu, sd), data = df$rating) %>%
#   dag_node("mu", "mu", rhs = int+design %*% coefs, child = "rating") %>%
#   dag_node("sd", "sd", rhs = cauchy(0,3,truncation = c(0, Inf)), child = "rating") %>%
#   dag_node("int", "int", rhs = normal(0,10), child = "mu") %>%
#   dag_node("coefs", "coefs", rhs = normal(0,10, dim=ncol(design)), child = "mu") %>%
#   dag_node("design", "design", data = design, child = "mu") %>%
#   dag_plate("Observations", "i", nodeLabels = c("rating","design","mu"))
```

# Example 3: Multiple Poisson regression

```{r}
# A multiple Bayesian linear regression model using the warpbreaks data.
# greta
data("warpbreaks")
X <- as_data(model.matrix(breaks ~ wool + tension, warpbreaks))
y <- as_data(warpbreaks$breaks)
int <- variable()
coefs <- normal(0, 5, dim = ncol(X) - 1)
beta <- c(int, coefs)
eta <- X %*% beta
distribution(y) <- poisson(exp(eta))
gretaModel <- model(int, coefs, eta)
draws <- mcmc(gretaModel)
tidyDrawsDF <- replaceLabels(draws) %>%
  as.matrix() %>%
  as_tibble() %>%
  addPriorGroups()
tidyDrawsDF %>% glimpse()
tidyDrawsDF %>%
  pull(param) %>%
  unique()
tidyDrawsDF %>%
  filter(param == "coefs[1,1]") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>%
  filter(param == "eta[13,1]") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>%
  filter(param == "int") %>%
  pull(value) %>%
  mean()

tidyDrawsDF <- bind_rows(
  draws$`11` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 1) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value"),
  draws$`12` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 2) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value"),
  draws$`13` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 3) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value"),
  draws$`14` %>% as.matrix() %>% as_tibble() %>% rowid_to_column() %>% mutate(chain = 4) %>%
    pivot_longer(cols = -c(rowid, chain), names_to = "param", values_to = "value")
)
#

# causact
XX <- model.matrix(breaks ~ wool + tension, warpbreaks)
yy <- warpbreaks$breaks
graph <- dag_create() %>%
  dag_node("yy", "yy", rhs = poisson(exp(eta)), data = yy) %>%
  dag_node("eta", "eta", rhs = XX %*% beta, child = "yy") %>%
  dag_node("beta", "beta", rhs = c(int, coefs), child = "eta") %>%
  dag_node("int", "int", rhs = normal(0, 1), child = "beta") %>% # variable() in causact it wants lower & upper
  dag_node("coefs", "coefs", rhs = normal(0, 5, dim = ncol(XX) - 1), child = "beta") %>%
  dag_node("XX", "XX", data = XX, child = "eta", keepAsDF = TRUE) %>%
  dag_plate("Observations", "i", nodeLabels = c("yy", "eta", "XX"))
graph %>% dag_render()
set.seed(2022)
drawsDF <- graph %>% dag_greta(mcmc = TRUE)
drawsDF %>% dagp_plot()
drawsDF %>% glimpse()
```

# Example 4: Multiple categorical regression

```{r}
# A multi-variable Bayesian categorical regression model using the iris data.

# greta
data(iris)
X <- as_data(cbind(1, iris[, 1:4]))
y <- model.matrix(~ Species - 1, iris)
P <- ncol(X)
K <- ncol(y)
beta <- normal(0, 5, dim = c(P, K - 1))
eta <- X %*% beta
prob <- imultilogit(eta)
distribution(y) <- categorical(prob)
gretaModel <- model(prob, eta, beta) # MODEL
draws <- mcmc(gretaModel)
tidyDrawsDF <- replaceLabels(draws) %>%
  as.matrix() %>%
  as_tibble() %>%
  addPriorGroups()
tidyDrawsDF %>% glimpse()
tidyDrawsDF %>%
  pull(param) %>%
  unique()
tidyDrawsDF %>%
  filter(param == "eta[90,1]") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>%
  filter(param == "eta[90,2]") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>%
  filter(param == "prob[90,1]") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>%
  filter(param == "prob[90,2]") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>%
  filter(param == "prob[90,3]") %>%
  pull(value) %>%
  mean()
tidyDrawsDF %>% glimpse()
tidyDrawsDF %>%
  pull(param) %>%
  unique()
tidyDrawsDF %>%
  filter(str_detect(param, "beta")) %>%
  group_by(param) %>%
  summarize(mn = mean(value))

# causact
XX <- cbind(1, iris[, 1:4]) %>% as.matrix()
yy <- model.matrix(~ Species - 1, iris)
P <- ncol(XX)
K <- ncol(yy)
graph <- dag_create() %>%
  dag_node("yy", "yy", rhs = categorical(prob), data = yy) %>%
  dag_node("prob", "prob", rhs = imultilogit(eta), child = "yy") %>%
  dag_node("eta", "eta", rhs = XX %*% beta, child = "prob") %>%
  dag_node("XX", "XX", data = XX, child = "eta") %>%
  dag_node("beta", "beta", rhs = normal(0, 5, dim = c(P, K - 1)), child = "eta") %>%
  dag_plate("Observations", "i", nodeLabels = c("yy", "XX", "eta", "prob"))
graph %>% dag_render()
set.seed(2022)
drawsDF <- graph %>% dag_greta(mcmc = TRUE)
drawsDF %>% dagp_plot()
drawsDF %>% glimpse()
```

I don't know why the graph %>% dag_render() yields extra nodes :(

# Example 5: Multiple linear regression with LASSO prior

```{r}
# A multi-variable Bayesian linear regression model using an exponential-normal prior for the coefficients.

# greta
data(attitude)
design <- as.matrix(attitude[, 2:7])
int <- normal(0, 10)
sd <- cauchy(0, 3, truncation = c(0, Inf))
tau <- exponential(0.5, dim = ncol(design))
coefs <- normal(0, tau)
mu <- int + design %*% coefs
distribution(attitude$rating) <- normal(mu, sd)
gretaModel <- model(int, sd, tau, coefs, mu)
draws <- mcmc(gretaModel)
tidyDrawsDF <- replaceLabels(draws) %>%
  as.matrix() %>%
  as_tibble() %>%
  addPriorGroups()
tidyDrawsDF %>% glimpse()
tidyDrawsDF %>%
  pull(param) %>%
  unique()
tidyDrawsDF %>%
  filter(str_detect(param, "coefs")) %>%
  group_by(param) %>%
  summarize(mn = mean(value), sd = sd(value))
tidyDrawsDF %>%
  filter(str_detect(param, "int")) %>%
  group_by(param) %>%
  summarize(mn = mean(value), sd = sd(value))
tidyDrawsDF %>%
  filter(str_detect(param, "sd")) %>%
  group_by(param) %>%
  summarize(mn = mean(value), sd = sd(value))
tidyDrawsDF %>%
  filter(str_detect(param, "tau")) %>%
  group_by(param) %>%
  summarize(mn = mean(value), sd = sd(value))
tidyDrawsDF %>%
  filter(str_detect(param, "mu")) %>%
  group_by(param) %>%
  summarize(mn = mean(value), sd = sd(value))

# causact
design <- as.matrix(attitude[, 2:7])
graph <- dag_create() %>%
  dag_node("rating", "rating", rhs = normal(mu, sd), data = attitude$rating) %>%
  dag_node("mu", "mu", rhs = int + design %*% coefs, child = "rating") %>%
  dag_node("sd", "sd", rhs = cauchy(0, 3, truncation = c(0, Inf)), child = "rating") %>%
  dag_node("coefs", "coefs", rhs = normal(0, tau, dim = ncol(design)), child = "mu") %>%
  dag_node("tau", "tau", rhs = exponential(0.5, dim = ncol(design)), child = "coefs") %>%
  dag_node("int", "int", rhs = normal(0, 10), child = "mu") %>%
  dag_node("design", "design", data = design, child = "mu", keepAsDF = TRUE) %>%
  dag_plate("Observations", "i", nodeLabels = c("rating", "design", "mu"))
graph %>% dag_render()
set.seed(2022)
drawsDF <- graph %>% dag_greta(mcmc = TRUE)
drawsDF %>% dagp_plot()
```



# Appendices

<details>

<summary>

References

</summary>

[greta](https://greta-stats.org/index.html)  
[causact](https://www.causact.com/)  

</details>

<details>

<summary>

SessionInfo

</summary>

```{r}
sessionInfo()
```

</details>
